Speculation in foreign exchange: noise or information?

### 2015-11-02 Knitr/ElsPaperMain.tex
Updated ElsPaper The main paper has been broken into Front and Main for the assessment.

\documentclass[preprint,12pt,authoryear]{elsarticle}
%\graphicspath{{./Figures/}} % Specifies the directory where pictures are stored
%\usepackage[dcucite]{harvard}
% \documentclass[final,1p,times,authoryear]{elsarticle}
%\documentclass[final,1p,times,twocolumn,authoryear]{elsarticle}
%\documentclass[final,3p,times,authoryear]{elsarticle}
%\documentclass[final,3p,times,twocolumn,authoryear]{elsarticle}
%\documentclass[final,5p,times,authoryear]{elsarticle}
%\documentclass[final,5p,times,twocolumn,authoryear]{elsarticle}
\usepackage{rotating}
\usepackage{amsmath}
\usepackage{setspace}
\usepackage{pdflscape}
\usepackage[flushleft]{threeparttable}
\usepackage{multirow}
\bibliographystyle{agsm}
\usepackage[colorlinks = true, citecolor = blue, linkcolor = blue]{hyperref}
%\hypersetup{urlcolor=blue, colorlinks=true} % Colors hyperlinks in blue - change to black if annoying
%\renewcommand[\harvardurl]{URL: \url}
\begin{document}
\begin{frontmatter}
\title{Speculation in foreign exchange: noise or information?}
\author{Rob Hayward\footnote{University of Brighton Business School, Lewes Road, Brighton, BN2 4AT; Telephone 01273 642586.  rh49@brighton.ac.uk.}}

\begin{abstract}
If foreign exchange prices are driven by speculation based on noise rather than information, extreme levels of speculative sentiment and speculative activity should lead to price reversals. Informed speculation is sustainable as it is part of the price discovery process. A unique set of option prices measure speculative sentiment while regulatory positions determine the weight of speculative activity.  An event study shows that extremes of speculative sentiment or speculative activity do not increase the probability of price reversal, though variance-measured risk increases.  Price activity after the extreme is close to a random walk. This supports the view that speculation is informed and part of the process of price-discovery rather than uninformed noise.      
\end{abstract}
\begin{keyword}
JEL classification: G12, G14, G15; event studies, information, noise-trading, speculation.
\end{keyword}

\end{frontmatter}
\section*{Introduction}
Ideas about the way that speculation affects financial markets span a broad range.  Towards one end is the view of \citet[p. 101]{Keynes1936} that speculation is a myopic, sentiment-driven activity, dominated by the desire to ``beat the gun'' or  ``outwit the crowd''. Speculation is a problem that will cause booms, reversals and will discourage fundamental, long-term investment. It is noise that emerges from the interaction of collective sentiment to create trends in asset prices.

@@

It is encouraging that two very different measures of speculative activity produce very similar results and that the results hold over different time frames: one week for the option data and one week for the regulatory position data.  It is less clear whether the results would be reproduced across markets for other financial assets. The results do not rule out uninformed speculation.  The Friedman effect from the Noise-trader model, whereby the informed speculators take advantage of the space created by price pressure from the uniformed, depends on there being \emph{noise-trader risk} or some limits to arbitrage.  The effect is not evident in this study.  However, as discussed in Footnote \ref{FX}, the foreign exchange market is a particularly large and liquid market, other asset classes may have alternative institutional features that mean a different relationship or balance between the informed and uninformed and therefore different results.

This is important because there is no unambiguous view of speculative activity and because speculation is an important issues in a number of policy debates.  Most notably, the \emph{Tobin Tax} would put a charge against financial transactions, particularly in foreign exchange.  Part of the argument depends on speculation being uninformed as the reduced activity that is caused by the tax would, it is argued, limit volatility and allow more long-term fundamental evaluations to become more prominent.  However, if speculation is informed and part of the price discovery process, this tax may have the opposite effect to what is envisaged.  These results suggest that it could reduce liquidity and increase volatility.

>>>>

\documentclass[preprint,12pt,authoryear]{elsarticle}
%\graphicspath{{./Figures/}} % Specifies the directory where pictures are stored
%\usepackage[dcucite]{harvard}
% \documentclass[final,1p,times,authoryear]{elsarticle}
%\documentclass[final,1p,times,twocolumn,authoryear]{elsarticle}
%\documentclass[final,3p,times,authoryear]{elsarticle}
%\documentclass[final,3p,times,twocolumn,authoryear]{elsarticle}
%\documentclass[final,5p,times,authoryear]{elsarticle}
%\documentclass[final,5p,times,twocolumn,authoryear]{elsarticle}
\usepackage{rotating}
\usepackage{amsmath}
%\usepackage{setspace}
\usepackage{pdflscape}
\usepackage[flushleft]{threeparttable}
\usepackage{multirow}
\bibliographystyle{agsm}
\usepackage[colorlinks = true, citecolor = blue, linkcolor = blue]{hyperref}
\begin{document}

\section*{Introduction}
Ideas about the way that speculation affects financial markets span a broad range.  Towards one end is the view of \citet[p. 101]{Keynes1936} that speculation is a myopic, sentiment-driven activity, dominated by the desire to ``beat the gun'' or  ``outwit the crowd''. Speculation is a problem that will cause booms, reversals and will discourage fundamental, long-term investment. It is noise that emerges from the interaction of collective sentiment to create trends in asset prices.

@@

It is encouraging that two very different measures of speculative activity produce very similar results and that the results hold over different time frames: one week for the option data and one week for the regulatory position data.  It is less clear whether the results would be reproduced across markets for other financial assets. The results do not rule out uninformed speculation.  The Friedman effect from the Noise-trader model, whereby the informed speculators take advantage of the space created by price pressure from the uniformed, depends on there being \emph{noise-trader risk} or some limits to arbitrage.  The effect is not evident in this study.  However, the foreign exchange market is a particularly large and liquid market, other asset classes may have alternative institutional features that mean a different relationship or balance between the informed and uninformed and therefore different results.

This is important because there is no unambiguous view of speculative activity and because speculation is an important issues in a number of policy debates.  Most notably, the \emph{Tobin Tax} would put a charge against financial transactions, particularly in foreign exchange.  Part of the argument depends on speculation being uninformed as the reduced activity that is caused by the tax would, it is argued, limit volatility and allow more long-term fundamental evaluations to become prominent.  However, if speculation is informed and part of the price discovery process, this tax may have the opposite effect to what is envisaged.  These results suggest that it could reduce liquidity and increase volatility.
